The Galactic Midplane is Not a Plane: Implications for dynamical analysis with {\em Gaia} data and beyond

### 2019-04-22 paper/ms.tex
Add Friske+ velocity variations, OOM midplane estimate

\subsection{LSR Variations} \label{ssec:lsr_var}
We also expect that the local standard of rest (LSR) should vary as a function
of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
the components of the LSR as a function of azimuth, but performing a best-fit
subtraction to correct for misalignment of the original coordinate system (as
in the previous section) is more involved. Since we find that the variation in
the LSR is less pronounced than for the midplane, and since the dominant
effect is limited to $J_r$ (Figure~\ref{fig:many_orbit_wrong_ref}), we defer
this calculation to future work.

>>>>

\subsection{Velocity Variations} \label{ssec:lsr_var}
We also expect that the local standard of rest (LSR) should vary as a function
of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
the components of the LSR as a function of azimuth, but performing a best-fit
subtraction to correct for misalignment of the original coordinate system (as
in the previous section) is more involved. Since we find that the variation in
the LSR is less pronounced than for the midplane, and since the dominant
effect is limited to $J_r$ (Figure~\ref{fig:many_orbit_wrong_ref}), we defer
this calculation to future work.

An interesting, qualitative calculation can be made based on the velocity
variations seen in e.g. \citet{2019arXiv190209569F}. The velocity variations
in this work (admittedly found with respect to $L_z$ or radius) are
$\sim5\text{--}10\,\kms$. The vertical frequency of the thin-disk and
thick-disk orbits are $\sim 0.09\,1/\Myr$ and $\sim0.06\,1/\Myr$,
respectively. By dimensional analysis, we therefore expect the midplane
offsets to be of order $\sim 57\text{--}170\,\pc$. We stress that this is a
rough calculation, but it shows that already we see evidence in the data from
velocities for midplane offsets on the order of what we see in both sets of
simulations.
